---
title: "Investigating Deepseek’s Influence on Public Sentiment and Tech Stock Performance"
subtitle: "Timeseries Final Project"
author: "Roopa Yashaswini Kotha"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
---

# Introduction

In January 2025, the artificial intelligence landscape witnessed a significant shift with the release of DeepSeek-R1 on January 20. This open-source large language model from China showcased reasoning capabilities comparable to leading models like OpenAI's o1, marking a pivotal moment in global AI competition. As a consequence, NVIDIA stock price declined drastically. However, the excitement was short-lived.

This report dives into the **cross-market ripple effects** of Deepseek's rise. We ask a central question:

*Can sudden bursts of attention around emerging AI technologies—like Deepseek—actually move the stock market, especially for companies like Nvidia and Microsoft that are at the heart of the AI ecosystem?*

**Data Overview**

**Twitter and Google News Activity**: Collected tweets and Google news headlines containing references to Deepseek, using them to derive daily tweet volumes and average sentiment scores. This reflects how often and how positively people were discussing Deepseek online.

**Stock Market Performance**: Tracked the daily closing prices and returns of two AI-adjacent tech titans: Nvidia (NVDA), a leader in GPU manufacturing and AI hardware, and Microsoft (MSFT), a major player in software and cloud infrastructure with strategic stakes in AI.

**Google Trends Search Data**: To understand broader public interest beyond Twitter, we also pulled Google search activity related to Deepseek, giving us a complementary view of curiosity and attention.

Aim is to test whether Deepseek’s moment in the spotlight had statistically significant influence on stock returns. Through correlation analysis, regression modeling, and hypothesis testing, we assess how real-time attention interacts with real-world markets.

# Imports

```{r, include=FALSE, warning= FALSE}
library(dplyr)
library(stringr)
library(textclean)
library(tm)
library(sentimentr)
library(lubridate)
library(quantmod)
library(ggplot2)
library(forecast)
library(scales)
library(gridExtra)
```

# Data Gathering

```{r, warning= FALSE}
#Data loading
dec_jan = read.csv('./dec-jan.csv')
jan_feb = read.csv('./jan-feb.csv')
feb_mar = read.csv('./feb-mar.csv')
mar_apr = read.csv('./mar-apr.csv')

dec_jan <- dec_jan %>% 
  select(title, published.date) %>% 
  rename(date = published.date) %>% 
  mutate(date = as.POSIXct(date, format="%a, %d %b %Y %H:%M:%S GMT", tz="GMT")) %>% 
  mutate(date = as.Date(date))

jan_feb <- jan_feb %>% 
  select(title, published.date) %>% 
  rename(date = published.date) %>%
  mutate(date = as.POSIXct(date, format="%a, %d %b %Y %H:%M:%S GMT", tz="GMT")) %>% 
  mutate(date = as.Date(date))

feb_mar <- feb_mar %>% 
  select(title, published.date) %>% 
  rename(date = published.date) %>% 
  mutate(date = as.POSIXct(date, format="%a, %d %b %Y %H:%M:%S GMT", tz="GMT")) %>% 
  mutate(date = as.Date(date))

mar_apr <- mar_apr %>% 
  select(title, published.date) %>% 
  rename(date = published.date) %>% 
  mutate(date = as.POSIXct(date, format="%a, %d %b %Y %H:%M:%S GMT", tz="GMT")) %>% 
  mutate(date = as.Date(date))

deepseek_df <- read.csv("./for_export_deepseek.csv")

trends_df <- read.csv("multiTimeline.csv")

trends_df$date <- as.Date(trends_df$date, format = "%d-%m-%Y")
trends_df$deepseek <- as.numeric(gsub("[^0-9.]", "", trends_df$deepseek))
```

## Data Cleaning

```{r, warning= FALSE}
#Data cleaning
deepseek_df <- deepseek_df %>% 
  filter(lang == 'en') %>% 
  select('createdAt', 'text')

deepseek_df$createdAt <- ymd_hms(deepseek_df$createdAt)

deepseek_df$createdAt <- as.Date(deepseek_df$createdAt)

start_date <- as.Date("2025-01-20", format="%Y-%m-%d")
end_date <- as.Date("2025-01-29", format="%Y-%m-%d")

# Filter the data frame
jan_df <- deepseek_df %>%
  filter(createdAt >= start_date & createdAt <= end_date) %>% 
  filter(grepl("deepseek", text, ignore.case = TRUE)) %>% 
  rename(date = createdAt, title = text)

combined_df <- bind_rows(dec_jan, jan_feb, jan_df, feb_mar, mar_apr)

clean_text <- function(text) {
  text %>%
    str_remove_all("https?://\\S+") %>%  # Remove URLs
    str_remove_all("[^\\w\\s]") %>%      # Remove special characters
    str_remove_all("\\d+") %>%           # Remove numbers
    #emoji_replace("") %>%                # Remove emojis
    removeWords(stopwords("en")) %>%     # Remove stopwords
    str_remove_all("[\r\n]+") %>%        # Remove new lines
    str_squish()                         # Remove extra spaces
}

combined_df$cleaned_text <- sapply(combined_df$title, clean_text)
```

```{r, warning= FALSE}
sentiment_scores <- sentiment(combined_df$cleaned_text)
combined_df$sentiment_scores = sentiment_scores$sentiment
```

```{r, warning= FALSE}
daily_tweet_data <- combined_df %>%
  group_by(date) %>%
  summarize(tweet_count = n(),
            min_sentiment = min(sentiment_scores),
            max_sentiment = max(sentiment_scores),
            avg_sentiment = mean(sentiment_scores))

start_date <- min(daily_tweet_data$date)
end_date   <- max(daily_tweet_data$date)

# Download data from Yahoo Finance. Stock data will be stored in objects NVDA and MSFT.
getSymbols(c("NVDA", "MSFT"), src = "yahoo", from = start_date, to = end_date)

# Extract closing prices and convert to data frames
nvda_df <- data.frame(date = index(NVDA), NVDA_close = as.numeric(Cl(NVDA)))
msft_df <- data.frame(date = index(MSFT), MSFT_close = as.numeric(Cl(MSFT)))
```

# Data Exploration and Visualisation

```{r, warning= FALSE}
summary_stats <- summary(combined_df$sentiment_scores)
print("Summary Statistics for Sentiment Scores:")
print(summary_stats)

# Compute quantiles to see where most values fall
quantiles <- quantile(combined_df$sentiment_scores, probs = c(0.05, 0.25, 0.5, 0.75, 0.95), na.rm = TRUE)
print("Quantiles (5%, 25%, 50%, 75%, 95%):")
print(quantiles)
```

```{r, warning= FALSE}
ggplot(combined_df, aes(x = sentiment_scores)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  labs(title = "Collected Deepseek Tweet Sentiment Is Largely Centered Near Neutral",
       x = "Sentiment Score", y = "Frequency") +
  theme_minimal()
```

Most Deepseek-related tweets and news collected have sentiment scores clustered near zero, indicating that the majority of public commentary is moderately balanced or neutral.

```{r, warning= FALSE}
ggplot(daily_tweet_data, aes(x = date, y = avg_sentiment)) +
  geom_line(color = "red") +
  labs(title = "Deepseek Tweet Sentiment Fluctuates Markedly Over Time",
       x = "Date",
       y = "Average Sentiment") +
  theme_minimal()

```

The sentiment scores vary considerably over the observed period, suggesting that public opinion on Deepseek shifts in response to changing news, events, launches and comments."

```{r, warning= FALSE}
# Plot 4: Sentiment Scores Over Time (Scatter Plot)
ggplot(combined_df, aes(x = date, y = sentiment_scores)) +
  geom_point(alpha = 0.5, color = "darkgreen") +
  labs(title = "Deepseek Sentiment Scores Spike Notably in January End Before Stabilizing",
       x = "Date",
       y = "Sentiment Score") +
  theme_minimal()
```

A sharp surge in sentiment during late January is evident, likely tied to a major Deepseek event (getting recognised by large audience and becoming a competitor for OpenAI), after which sentiment levels stabilize, indicating a temporary reaction.

```{r, warning= FALSE}
# Plot the average sentiment for all tweets by date (with low volume markers possibly flagged)
ggplot(daily_tweet_data, aes(x = date, y = avg_sentiment, size = tweet_count)) +
  geom_point(color = "blue") +
  geom_line(aes(group = 1), color = "blue", alpha = 0.7) +
  labs(title = "Deepseek’s Daily Sentiment Fluctuates with Surges in Tweet Volume",
       subtitle = "Point size represents tweet count per date",
       x = "Date",
       y = "Average Sentiment Score") +
  theme_minimal()
```

Days with higher tweet volume are associated with greater volatility in sentiment, implying that intense media attention may trigger stronger positive or negative public reactions.

```{r, warning= FALSE}
key_events <- data.frame(
  event_date = as.Date(c("2025-01-22", "2025-01-23", "2025-01-25", "2025-01-27", "2025-01-29", "2025-02-04")),
  event_type = c("LLM Launch", "Recognition", "Recognition", "Competition", "Notable Comment", "Controversy"),
  event_detail = c("Large language model launch",
                   "Attention for its affordability",
                   "Rapid advancements in AI",
                   "Dethroning ChatGPT",
                   "Serious threat",
                   "OpenAI accuses copying"),
  # Assign a manual y position for each event so they don't overlap
  y_position = c(40000, 38000, 36000, 34000, 32000, 30000)  
)

```

```{r, warning= FALSE}
daily_tweet_data <- daily_tweet_data %>%
  arrange(date) %>%
  mutate(
    tweet_count_change = tweet_count - lag(tweet_count),
    sentiment_change = avg_sentiment - lag(avg_sentiment)
  )

print("Tweet Data Summary:")
print(summary(daily_tweet_data))
```

```{r, warning= FALSE}
merged_df <- daily_tweet_data %>%
  full_join(nvda_df, by = "date") %>%
  full_join(msft_df, by = "date") %>%
  left_join(trends_df, by = "date") %>% 
  arrange(date)

merged_df <- merged_df %>%
  arrange(date) %>%
  mutate(
    tweet_count = na.locf(tweet_count, na.rm = FALSE),
    min_sentiment = na.locf(min_sentiment, na.rm = FALSE),
    max_sentiment = na.locf(max_sentiment, na.rm = FALSE),
    avg_sentiment = na.locf(avg_sentiment, na.rm = FALSE),
    tweet_count_change = na.locf(tweet_count_change, na.rm = FALSE),
    sentiment_change = na.locf(sentiment_change, na.rm = FALSE),
    NVDA_return = (NVDA_close - lag(NVDA_close)) / lag(NVDA_close),
    MSFT_return = (MSFT_close - lag(MSFT_close)) / lag(MSFT_close)
  )

max_price <- max(merged_df$NVDA_close, na.rm = TRUE)
min_price <- min(merged_df$NVDA_close, na.rm = TRUE)
range_price <- max_price - min_price
```

```{r, warning= FALSE}
print("Merged Stock Data Summary:")
print(summary(merged_df))
```

```{r, warning= FALSE}
merged_df <- merged_df %>%
  mutate(sentiment_scaled = (avg_sentiment - min(avg_sentiment, na.rm = TRUE)) /
           diff(range(avg_sentiment, na.rm = TRUE)) * range_price + min_price)
```

```{r, warning= FALSE}
quantmod::chartSeries(NVDA, type="candlesticks", subset="2024-12-26::2025-04-02", TA=NULL)
```

```{r, warning= FALSE}
quantmod::chartSeries(MSFT, type="candlesticks", subset="2024-12-26::2025-04-02", TA=NULL)
```

```{r, warning= FALSE}
library(lubridate)

ggplot(merged_df, aes(x = date, y = tweet_count)) +
  # Add shaded region
  geom_rect(aes(xmin = as.Date("2025-01-22"), xmax = as.Date("2025-02-04"),
                ymin = -Inf, ymax = Inf),
            fill = "red", alpha = 0.01, inherit.aes = FALSE) +
  # Add tweet count line
  geom_line(color = "steelblue", size = 1) +
  labs(
    title = "Deepseek Tweet Count with Key Events",
    x = "Date",
    y = "Tweet Count"
  ) +
  theme_minimal()
```

```{r, warning= FALSE}
y_top <- max(merged_df$avg_sentiment, na.rm = TRUE)
y_label <- y_top - 0.3

p2 <- merged_df %>% 
  mutate(
    NVDA_close = na.locf(NVDA_close, na.rm = FALSE),
    MSFT_close = na.locf(MSFT_close, na.rm = FALSE)
  ) %>%
  mutate(
    nvda_scaled = scale(NVDA_close)[,1],
    msft_scaled = scale(MSFT_close)[,1]
  ) %>% 
  filter(date >= dmy("17/01/2025") & date <= dmy("10/02/2025")) %>% 
  ggplot(aes(x = date, y = avg_sentiment)) +
  geom_line(aes(y = avg_sentiment), color = "darkgreen", size = 1) +
  geom_line(aes(y = nvda_scaled), color = "purple", size = 0.8) +
  geom_line(aes(y = msft_scaled), color = "blue", size = 0.8) +
  labs(title = "Impact of Deepseek Events on Public Sentiment and Tech Stock Prices",
       x = "Date",
       y = "Average Sentiment") +
  scale_color_manual(values = c(
    "avg_sentiment" = "darkgreen",
    "nvda_scaled" = "purple",
    "msft_scaled" = "blue"
  )) +
  theme_minimal()

for(i in 1:nrow(key_events)) {
  p2 <- p2 +
    geom_vline(xintercept = as.numeric(key_events$event_date[i]),
               linetype = "dashed", color = "red", alpha = 0.7) +
    annotate("text",
             x = key_events$event_date[i],
             y = y_label,  # adjust as needed
             label = paste(key_events$event_type[i], ":", key_events$event_detail[i]),
             angle = 90, vjust = 0, size = 3, color = "red")
}


print(p2)
```

The graph reveals that while public sentiment around Deepseek remained moderately positive, key media events triggered noticeable volatility in NVDA and MSFT stock prices—particularly around the "Dethroning ChatGPT" and "OpenAI Accusation" episodes, suggesting market sensitivity to high-stakes AI narratives.

```{r, warning= FALSE}
ggplot(merged_df, aes(x = date)) +
  geom_line(aes(y = NVDA_close, color = "NVDA Close"), size = 1) +
  geom_line(aes(y = sentiment_scaled, color = "Avg Sentiment (Scaled)"), linetype = "dashed", size = 1) +
  scale_y_continuous(
    name = "NVDA Closing Price",
    sec.axis = sec_axis(~ (. - min_price) / range_price * diff(range(merged_df$avg_sentiment, na.rm=TRUE)) +
                          min(merged_df$avg_sentiment, na.rm = TRUE),
                        name = "Avg Sentiment")
  ) +
  labs(title = "NVDA Stock Price and Scaled Deepseek Tweet Sentiment Fluctuate Over Time",
       x = "Date") +
  scale_color_manual(values = c("NVDA Close" = "purple", "Avg Sentiment (Scaled)" = "orange")) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r, warning= FALSE}
# 4.3 Scatter Plot: Tweet Sentiment vs NVDA Daily Return
# This helps in exploring any association between tweet mood and stock performance.
ggplot(merged_df, aes(x = avg_sentiment, y = NVDA_return)) +
  geom_point(alpha = 0.7, color = "darkblue") +
  geom_smooth(method = "lm", se = TRUE, color = "orange") +
  labs(title = "Deepseek Tweet Sentiment Shows a Weak Negative Correlation with NVDA Daily Returns",
       x = "Average Tweet Sentiment", y = "NVDA Daily Return") +
  theme_minimal()
```

```{r, warning= FALSE}
if(nrow(nvda_df) > 30) {
  # Convert NVDA closing prices to a time series object.
  # Frequency set to 5 assuming ~5 trading days per week.
  nvda_ts <- ts(nvda_df$NVDA_close, frequency = 5)
  nvda_decomp <- stl(nvda_ts, s.window = "periodic")
  
  # Plot the decomposition
  plot(nvda_decomp, main = "Seasonal Decomposition of NVDA Stock Prices")
} else {
  message("Not enough NVDA stock data for seasonal decomposition.")
}
```

The seasonal decomposition of NVDA stock prices reveals a strong and regular seasonal component, alongside a gradually shifting trend. The remainder (residual) indicates periods of irregular volatility likely driven by news or external shocks beyond normal patterns.

```{r, warning= FALSE}
if(nrow(msft_df) > 30) {
  # Using a frequency of 5 for weekly patterns (assuming trading days)
  msft_ts <- ts(msft_df$MSFT_close, frequency = 5)
  msft_decomp <- stl(msft_ts, s.window = "periodic")
  
  plot(msft_decomp, main = "Seasonal Decomposition of MSFT Stock Prices")
} else {
  p_decomp <- ggplot() +
    geom_text(aes(0,0,label = "Not enough data for seasonal decomposition"), hjust = 0.5, vjust = 0.5) +
    theme_void()
}
```

The decomposition of MSFT stock prices reveals a clear cyclical seasonal pattern and a relatively stable trend with mild downward shifts. The low variability in the remainder component suggests MSFT prices were less influenced by unexpected shocks during this period compared to NVDA.

# Analysis

## Hypothesis Formation and Quantitative Analysis Techniques

**Null Hypothesis (H0):** There is no association between Deepseek-related media/social media attention and the daily stock returns of Nvidia and Microsoft.In other words, any observed fluctuations in stock prices would be purely coincidental and unrelated to the level of attention Deepseek receives.

**Alternate Hypothesis {H1):** Increases in Deepseek-related media and social media attention (measured by tweet volume and average sentiment) are associated with significant movements in Nvidia and Microsoft stock prices.

## Correlation Analysis

We'll compute correlation coefficients between tweet-level metrics (tweet_count and avg_sentiment) and daily stock returns.

```{r, warning= FALSE}
# Basic Correlations
cor_nvda_count <- cor(merged_df$tweet_count, merged_df$NVDA_return, use = "complete.obs")
cor_nvda_sentiment <- cor(merged_df$avg_sentiment, merged_df$NVDA_return, use = "complete.obs")
cor_msft_count <- cor(merged_df$tweet_count, merged_df$MSFT_return, use = "complete.obs")
cor_msft_sentiment <- cor(merged_df$avg_sentiment, merged_df$MSFT_return, use = "complete.obs")

cat("Correlation between tweet count and NVDA return:", cor_nvda_count, "\n")
cat("Correlation between average sentiment and NVDA return:", cor_nvda_sentiment, "\n")
cat("Correlation between tweet count and MSFT return:", cor_msft_count, "\n")
cat("Correlation between average sentiment and MSFT return:", cor_msft_sentiment, "\n")
```

```{r, warning= FALSE}
cor_trend_nvda <- cor(merged_df$deepseek, merged_df$NVDA_return, use = "complete.obs")
cor_trend_msft <- cor(merged_df$deepseek, merged_df$MSFT_return, use = "complete.obs")
cor_trend_tweet <- cor(merged_df$deepseek, merged_df$tweet_count, use = "complete.obs")
cor_trend_sent  <- cor(merged_df$deepseek, merged_df$avg_sentiment, use = "complete.obs")

cat("Correlation between Google Trends and NVDA_return:", cor_trend_nvda, "\n")
cat("Correlation between Google Trends and MSFT_return:", cor_trend_msft, "\n")
cat("Correlation between Google Trends and tweet_count:", cor_trend_tweet, "\n")
cat("Correlation between Google Trends and avg_sentiment:", cor_trend_sent, "\n")
```

```{r, warning= FALSE}
# Lagged Correlations
# Create lagged versions of tweet_count and avg_sentiment to see if previous-day media attention predicts current day returns.
merged_df <- merged_df %>%
  arrange(date) %>%
  mutate(tweet_count_lag1 = lag(tweet_count, 1),
         avg_sentiment_lag1 = lag(avg_sentiment, 1))

cor_nvda_count_lag <- cor(merged_df$tweet_count_lag1, merged_df$NVDA_return, use = "complete.obs")
cor_nvda_sentiment_lag <- cor(merged_df$avg_sentiment_lag1, merged_df$NVDA_return, use = "complete.obs")

cat("Lagged (1 day) correlation between tweet count and NVDA return:", cor_nvda_count_lag, "\n")
cat("Lagged (1 day) correlation between avg sentiment and NVDA return:", cor_nvda_sentiment_lag, "\n")

cor_msft_count_lag <- cor(merged_df$tweet_count_lag1, merged_df$MSFT_return, use = "complete.obs")
cor_msft_sentiment_lag <- cor(merged_df$avg_sentiment_lag1, merged_df$MSFT_return, use = "complete.obs")

cat("Lagged (1 day) correlation between tweet count and NVDA return:", cor_msft_count_lag, "\n")
cat("Lagged (1 day) correlation between avg sentiment and NVDA return:", cor_msft_sentiment_lag, "\n")
```

```{r, warning= FALSE}
test_sentiment_nvda <- cor.test(merged_df$avg_sentiment, merged_df$NVDA_return, method = "pearson")
cat("Significance test for avg sentiment and NVDA return:\n")
print(test_sentiment_nvda)

test_sentiment_msft <- cor.test(merged_df$avg_sentiment, merged_df$MSFT_return, method = "pearson")
cat("Significance test for avg sentiment and MSFT return:\n")
print(test_sentiment_msft)

test_count_nvda_lag <- cor.test(merged_df$tweet_count_lag1, merged_df$NVDA_return, method = "pearson")
cat("Significance test for lagged tweet count and NVDA return:\n")
print(test_count_nvda_lag)

test_sentiment_nvda_lag <- cor.test(merged_df$avg_sentiment_lag1, merged_df$NVDA_return, method = "pearson")
cat("Significance test for lagged avg sentiment and NVDA return:\n")
print(test_sentiment_nvda_lag)
```

Correlation analysis between tweet metrics, Google Trends, and stock returns provided the following insights: For NVDA, the tweet count showed a moderate same-day positive correlation, whereas average sentiment had a weak negative association.

For MSFT, the correlations were weaker, suggesting that Deepseek-related media may influence NVDA more strongly. However, when non-neutral tweets were isolated, the signals became even more muted. These results suggest that while there might be an immediate market reaction to increased media activity, the effect may dissipate quickly or be masked by other factors.

## Regression Analysis

We build linear regression models to see if Deepseek media metrics can predict stock returns. Here, we treat NVDA_return (and then MSFT_return) as the dependent variable.

```{r, warning= FALSE}
# Model 1: Using same-day metrics
lm_nvda <- lm(NVDA_return ~ tweet_count + avg_sentiment, data = merged_df)
cat("Regression Model for NVDA_return (same day metrics):\n")
print(summary(lm_nvda))

lm_msft <- lm(MSFT_return ~ tweet_count + avg_sentiment, data = merged_df)
cat("Regression Model for MSFT_return (same day metrics):\n")
print(summary(lm_msft))

# Model 2: Using lagged predictors (e.g., previous day's tweet count and sentiment)
lm_nvda_lag <- lm(NVDA_return ~ tweet_count_lag1 + avg_sentiment_lag1, data = merged_df)
cat("Regression Model for NVDA_return (lagged predictors):\n")
print(summary(lm_nvda_lag))

lm_msft_lag <- lm(MSFT_return ~ tweet_count_lag1 + avg_sentiment_lag1, data = merged_df)
cat("Regression Model for NVDA_return (lagged predictors):\n")
print(summary(lm_msft_lag))
```

```{r, warning= FALSE}
lm_nvda_full <- lm(NVDA_return ~ tweet_count + avg_sentiment + deepseek, data = merged_df)
lm_msft_full <- lm(MSFT_return ~ tweet_count + avg_sentiment + deepseek, data = merged_df)

print(summary(lm_nvda_full))
print(summary(lm_msft_full))
```

Regression models incorporating tweet volume, average sentiment, and Google Trends data indicate that for NVDA, tweet volume is a statistically significant predictor of daily returns. In contrast, average sentiment does not achieve significance, and for MSFT, neither predictor was significant. This finding implies that the market’s response to Deepseek events might be more pronounced for some stocks (like NVDA) compared to others, or that simple linear models might not fully capture the dynamics at play

## Hypothesis Testing

We test if periods of high media attention are significantly different in terms of stock performance.

```{r, warning= FALSE}
# Define "high media attention" as days when tweet_count exceeds the median tweet count.
merged_df <- merged_df %>%
  mutate(high_media = tweet_count > median(tweet_count, na.rm = TRUE))

# T-test for NVDA returns: Comparing days with high vs. low media attention
t_test_nvda <- t.test(NVDA_return ~ high_media, data = merged_df)
cat("T-test for NVDA returns (high vs. low media attention):\n")
print(t_test_nvda)

# T-test for MSFT returns: Comparing days with high vs. low media attention
t_test_msft <- t.test(MSFT_return ~ high_media, data = merged_df)
cat("T-test for MSFT returns (high vs. low media attention):\n")
print(t_test_msft)
```

We performed t-tests to compare stock returns on days of high versus low media attention (as indicated by tweet counts and search interest). For NVDA, the test suggested a trend toward higher returns on high media attention days, but the result was only marginally significant. In the case of MSFT, no significant difference was found. These outcomes provide partial support for our hypothesis that Deepseek-related media attention affects stock performance, particularly for NVDA, albeit not consistently across all metrics.

```{r, warning= FALSE}
library(lmtest)

# Granger causality test: Do past tweet counts help predict NVDA_return?
granger_nvda <- grangertest(NVDA_return ~ tweet_count, order = 1, data = merged_df)
cat("Granger Causality Test (NVDA_return ~ tweet_count):\n")
print(granger_nvda)

# Granger causality test: Do past average sentiment values help predict NVDA_return?
granger_nvda_sentiment <- grangertest(NVDA_return ~ avg_sentiment, order = 1, data = merged_df)
cat("Granger Causality Test (NVDA_return ~ avg_sentiment):\n")
print(granger_nvda_sentiment)
```

The Granger causality analysis does not support the claim that historical tweet count or sentiment, when lagged by one day, reliably predicts NVDA’s daily stock returns. Although same-day correlations hinted at a moderate relationship (specifically for tweet count), the lack of statistical significance in the Granger tests (F = 2.05, p = 0.12 for tweet count and F = 1.45, p = 0.24 for average sentiment) suggests that any delayed effect of Deepseek-related media may be minimal or obscured by other market factors. Future analyses might explore different lag structures or incorporate additional variables to further investigate these dynamics.

## Text Analysis

```{r, warning= FALSE}
# -------------------------------
# 5. Text Analysis Insights Visualization
# -------------------------------
# Visualize the distribution of the daily average sentiment to better understand its behavior.
ggplot(merged_df, aes(x = avg_sentiment)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Daily Average Tweet Sentiment", 
       x = "Average Sentiment", y = "Frequency") +
  theme_minimal()
```

```{r, warning= FALSE}
trend_ts <- ts(trends_df$deepseek, start = c(year(min(trends_df$date)), yday(min(trends_df$date))), frequency = 365)

# (b) Use auto.arima to fit a model and forecast future trends.
trend_fit <- auto.arima(trend_ts)
trend_forecast <- forecast(trend_fit, h = 30)  # Forecast for the next 30 days
plot(trend_forecast, main = "Forecast for Google Trends Interest for Deepseek")


```

The Google Trends data for Deepseek demonstrates clear peaks in search interest that align with major news events, reinforcing the connection between public awareness and market reactions.

# Results

**Correlations:** The observed correlation between tweet volume and NVDA returns supports the idea that surges in Deepseek media attention may be linked with market movement. However, the weak or non-significant correlations for sentiment and for MSFT indicate that this relationship may be complex and not uniformly strong.

**Regression Findings:** Regression models reinforce that tweet volume has some explanatory power for NVDA returns. In contrast, average sentiment and similar predictors for MSFT did not show significant effects. This suggests that while media intensity might influence some stocks, the sentiment tone alone may be insufficient to drive market reactions reliably.

**Hypothesis Testing:** The t-tests comparing high versus low media attention days provided mixed evidence—there is an indication of an effect for NVDA, but not for MSFT. These mixed results hint that other factors, such as overall market sentiment and unrelated technological news, may also play crucial roles.

**Overall Implications:** Our analysis indicates that Deepseek-related media attention can have a measurable impact on stock performance—particularly for Nvidia—during key event periods. At the same time, the effects are not strongly consistent across all metrics, underscoring the need for more sophisticated models or longer-term analyses.

Future work might explore non-linear dynamics, additional lags, or incorporate broader market data to better isolate the effect of Deepseek on stock returns.

# Conclusion

Deepseek attention may influence Nvidia—but not Microsoft—and even for Nvidia, the effect is modest and primarily driven by the intensity of attention rather than its tone. Investors should be cautious in interpreting media buzz as a trading signal, especially when sentiment is not backed by strong fundamentals or broader market trends.


